---
title: "Delphine_Assignment5.Rmd"
output: html_notebook
---
# load packages
```{r}
library(ggthemes)
library(rstatix)
library(ggpubr)
library(tidyverse)
```
# import data 
```{r}
assignment5=read_csv("assignment5_data.csv")
nrow(assignment5)
ncol(assignment5)
colnames(assignment5)
```
# subject-level means
```{r}

subject_accuracy = assignment5 %>%
  group_by(subject, prime_condition)%>%
  summarise(mean=mean(accuracy))


subject_rt = assignment5%>%
  group_by(subject, prime_condition)%>%
  summarise(mean=mean(response_RT))
  
```
subject_accuracy is in long format, and subject_rt is also in long format

# long to wide conversion
```{r}
subject_accuracy_wide = subject_accuracy%>%
  pivot_wider(names_from = prime_condition, values_from = mean)
```
# wide to long
```{r}
subject_accuracy_long = subject_accuracy_wide%>%
  pivot_longer(names_to = "prime condition", cols = both:unrelated)
```

subject_accuracy contains the same information as that contained in subject_accuracy_long

# t test
```{r}
t.test(subject_accuracy_wide$phonological, subject_accuracy_wide$semantic, paired = TRUE)
```
we are concluding that they are not equal / there is a significant difference between semantic prime condition and phonological, because the p-value found is less than .025 (which is half of .05 because the test is two tailed) which we are using as each tail value.

#histogram
```{r}
subject_accuracy_wide = subject_accuracy_wide %>%
  mutate(diff = semantic - phonological)

subject_accuracy_wide = subject_accuracy_wide %>% ungroup()
```
```{r}
hist(subject_accuracy_wide$diff)
ggqqplot(subject_accuracy_wide, "diff")

```
# t test manual
```{r}
x_bar = mean(subject_accuracy_wide$diff)
sd_diff = sd(subject_accuracy_wide$diff)
t = x_bar/(sd_diff/sqrt(nrow(subject_accuracy_wide)))

df = nrow(subject_accuracy_wide)-1

p_value = 2*(1-pt(t, df))

t.test(subject_accuracy_wide$phonological, subject_accuracy_wide$semantic, var.equal = TRUE, paired=TRUE)
       

```

# t test outliers
```{r}
subject_accuracy_wide = subject_accuracy_wide %>% ungroup()

outliers = subject_accuracy_wide %>% identify_outliers(diff)

outlier_subs = outliers %>% pull(subject)

subject_wide_new = subject_accuracy_wide %>% filter(!subject %in% outlier_subs)
```
# test normality 
```{r}
ggqqplot(subject_accuracy_wide, "diff")

subject_accuracy_wide%>%
  shapiro_test(diff)
```
the normality assumption is satisfied- the test of assumptions does not change my conclusion - the t test result and individual results are grouped closely to the trend line. there is a very small p value and we can conclude that both phonological and semantic cues help people retrieve words from memory (both low and high means are correlated with semantic and phonological cues).

# plot RTs
```{r}
mean_rt = subject_accuracy %>%
  group_by(prime_condition)%>%
summarise(mean_rt=mean(mean),
          sd_rt=sd(mean))
```
```


